---
title: "TFcheckpoint Dataset visualization using UMAP"
author: "[John Zobolas](https://github.com/bblodfon)"
date: "Last updated: `r format(Sys.time(), '%d %B, %Y')`"
description: "Description"
url: 'https\://druglogics.github.io/tfcheckpoint-umap/'
github-repo: "druglogics/tfcheckpoint-umap"
bibliography: references.bib
link-citations: true
site: bookdown::bookdown_site
---

# Input {-}

Loading libraries:
```{r Load libraries, message = FALSE}
library(xfun)
```

# Intro {-}

TFcheckpoint [@Chawla2013] is a curated database with sequence-specific DNA-binding RNA polymerase II transcription-factor proteins (DbTF).

The form of the dataset is as follows: a total of $4705$ proteins, each one represented by a row, are enriched with GO annotations (a total of $8621$ GO terms [@Carbon2019]) that indicate the presence ($1$) or absence ($0$) of a particular molecular function. 
The dataset is provided as a [tab-delimited file](https://github.com/druglogics/tfcheckpoint-umap/blob/main/data/genes2go_result_tfcheckpoint2_data.txt).

In this analysis we present a 2D visualization of this dataset using the non-linear dimension reduction method UMAP [@McInnes2018a].
This method will reduce the size of the dataset from $(4705 \times 8621)$ to $(4705 \times 2)$ while efficiently identifying groups of proteins with the same molecular characterization. 
These groups are indicated by varying size clusters in the 2D plane.

# Analysis {-}

:::{.blue-box}
See script [umap.R](https://github.com/druglogics/tfcheckpoint-umap/blob/main/scripts/umap.R) for more details about the presented analysis.
:::

# R session info {-}

```{r session info, comment="", class.source = 'fold-show'}
xfun::session_info()
```

# References {-}
